---
title: "US Crime-Rate Bias Determiner"
author: "trumayanvoid"
date: "5/31/2020"
output: slidy_presentation
runtime: shiny
---
```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = FALSE)
```

## Purpose

Due to the highly-diverse population of United States combined with historicalism, national laws and policies, and culture, there is a high prevalence of hate crimes in the country.

Dataset used was from 2018 dataset from  https://ucr.fbi.gov/hate-crime/2018/downloads/hate-crime-statistics-2018-tables.zip. 

This application calculates the probability that you will be the target for a hate crime based on your several features. Also, this provides rankings for the number of victims and incidences per group/feature in 2018.


## Functions and Tools

1.Probability Determination based on Features

```{r, echo=FALSE}

shinyUI(fluidPage(
   sidebarLayout(
 sidebarPanel(selectInput("group1", "Closest Race/Ethnicity/Ancestry",
                        c("White"= "white", "Black or African American"= "black",
                          "American Indian or Alaska Native"= "native_a", "Asian"= "asian",
                          "Native Hawaiian or Other Pacific Islander"= "native_h",
                          "Multiple Races or Group"= "multiple_race", "Arab"= "arab", 
                          "Hispanic or Latino"= "hispanic", "Other Race/Ethnicity/Ancestry"= "others_race")), 

selectInput("group2", "Religious Affiliation/ Belief System", c("Jewish"= "jew", "Catholic"= "catholic", "Protestant" = "protestant", "Islamic" = "islam", "Other Religion"= "others_religion", 'Multiple Religions/Groups'= "multiple_religion","Mormon"= "mormon", "Jehova's Witness"= "jehova", "Eastern Orthodox (Russian, Greek, Other)"= "orthodox", "Other Christian"= "christian","Buddhist"= "buddhist", "Hindu"= "hindu", "Sikh" = "sikh", "Atheism/Agnoticism" = "aa", "None of the Above"= "none")),

selectInput("group3", "Sexual Orientation", c("Gay (Male)"= "gay", "Lesbian"= "lesbian", "Lesbian, Gay, Bisexual, Transgender (Mixed Group)" = "mixed_so", "Heterosexual"= "hetero", "Bisexual"= "bisexual")),

selectInput("group4", "Disability", c("Physical"= "physical", "Mental"= "mental", "None of the Above"= "none")),
            
selectInput("group5", "Gender", c("Male"= "male", "Female"= "female", "None of the Above"= "none")),
            
selectInput("group6", "Gender Identity", c("Transgender"= "trans", "Gender Non-Conforming"= "non", "None of the Above"= "none"))),

mainPanel(plotOutput("histogram1"), h4("Probability:"), textOutput("histogram2"))

)))

library(dplyr)
library(ggplot2)
library(tidyverse)

read.csv("Hate.csv") -> x
apply(x, 2, as.numeric) -> x
as.data.frame(t(x[,2:ncol(x)])) -> x
mutate(x, Groups = rownames(x)) -> x
colnames(x) = c("Incident", "Victims", "Victims/Incidents", "Groups")
x -> hate
sum(hate$Incident) -> total_incidents
sum(hate$Victims) -> total_victims

modelhate <- reactive({
        hateinput1 <- filter(hate, Groups %in% input$group1) 
        hateinput2 <- filter(hate, Groups %in% input$group2)
        hateinput3 <- filter(hate, Groups %in% input$group3)
        hateinput4 <- filter(hate, Groups %in% input$group4)
        hateinput5 <- filter(hate, Groups %in% input$group5)
        hateinput6 <- filter(hate, Groups %in% input$group6)
        totalhate <-rbind(hateinput1, hateinput2, hateinput3, hateinput4, hateinput5, hateinput6) %>%
        mutate(row= row_number())
        colorRampPalette(c("darkgreen", "yellowgreen")) -> c
        c(nrow(totalhate)) -> c
        ggplot(data= totalhate, aes(x= reorder(Groups,row), y= Victims)) + geom_col(fill=c) + ylim(range(0, 2500)) + labs(x= "Group", y= "Victims in 7,036 incidences of hate crime")  
            
    })

   modelbar <- reactive({
        hateinput1 <- filter(hate, Groups %in% input$group1) 
        hateinput2 <- filter(hate, Groups %in% input$group2)
        hateinput3 <- filter(hate, Groups %in% input$group3)
        hateinput4 <- filter(hate, Groups %in% input$group4)
        hateinput5 <- filter(hate, Groups %in% input$group5)
        hateinput6 <- filter(hate, Groups %in% input$group6)
        totalhate2 <-rbind(hateinput1, hateinput2, hateinput3, hateinput4, hateinput5, hateinput6) 
        totalhate2 <- totalhate2 %>% mutate(prop= (Victims/total_victims*100)) %>% select(Groups, prop) %>% summarize(sum(prop))
        paste(signif(totalhate2[1,1],4), "%", sep = "")
        
        
        
    })
output$histogram2 <- renderText({modelbar()})
output$histogram1 <- renderPlot({modelhate()})
```

## 2. Ranking of Group/Features by Number of Victims and Incidences against them
```{r, echo=FALSE}
sidebarPanel(checkboxInput("check", "Show Rankings", value = TRUE),

conditionalPanel(condition= "input.check == 1", selectInput("order", "Ranking order", c("Most Vulnerable to Least Vulnerable"= "dec", "Least Vulnerable to Most Vulnerable"= "inc")), sliderInput("rankslider", "Number of Groups to Show", 0,34, 10))
)

mainPanel(tabsetPanel(type= "tabs", tabPanel("Ranking by Incident", plotOutput("rank")),tabPanel("Ranking by Number of Victims", plotOutput("rank2"))))


modelrank <- reactive({
        hate <- hate %>% mutate(full = c("White", "Black/African American", "Asian", "Native Hawaiian/Pacific Islander", 
                                         "Multiple Races/Group", "Arab", "Hispanic", "Latino", "Other Race", "Jewish",
                                         "Catholic", "Protestant", "Muslim", "Other Religion", "Multiple Religions", "Mormon",
                                         "Jehova's Witness", "Eastern Orthodox", "Other Christian", "Buddhist", "Hindu", 
                                         "Sikh", "Atheism/Agnosticism/etc.", "Gay(Male)", "Lesbian", "LGBT(Mixed Group)",
                                         "Heterosexual", "Bisexual", "Physically Disabled", "Mentally Disabled", "Male", "Female", "Transgender", "Gender Non-Conforming"))
           
        if (input$check == TRUE) {ifelse(input$order=="dec", (hate<- arrange(hate, desc(Incident), desc(Victims), full)), (hate<- arrange(hate, Incident, Victims, full)) )}
        else {stop()}
        
        hate[1:input$rankslider,] -> hate
        mutate(hate, row= row_number()) -> hate
        colorRampPalette(c("red","yellow"))-> c
        c(input$rankslider) -> c
        ggplot(hate, aes(y= reorder(full, row), x= Incident)) + geom_col(fill= c) + labs(y= "Groups", x= "Number of Hate Crime Incidents Against A Group")
     
        
       
    })
    
    
    modelrank2 <- reactive({
        hate <- hate %>% mutate(full = c("White", "Black/African American", "Asian", "Native Hawaiian/Pacific Islander", 
                                         "Multiple Races/Group", "Arab", "Hispanic", "Latino", "Other Race", "Jewish",
                                         "Catholic", "Protestant", "Muslim", "Other Religion", "Multiple Religions", "Mormon",
                                         "Jehova's Witness", "Eastern Orthodox", "Other Christian", "Buddhist", "Hindu", 
                                         "Sikh", "Atheism/Agnosticism/etc.", "Gay(Male)", "Lesbian", "LGBT(Mixed Group)",
                                         "Heterosexual", "Bisexual", "Physically Disabled", "Mentally Disabled", "Male", "Female", "Transgender", "Gender Non-Conforming"))
        
        if (input$check == TRUE) {ifelse(input$order=="dec", (hate<- arrange(hate, desc(Victims), desc(Incident), full)), (hate<- arrange(hate, Victims, Incident, full)) )}
        else {stop()}
        
        hate[1:input$rankslider,] -> hate
        mutate(hate, row= row_number()) -> hate
        colorRampPalette(c("red","yellow"))-> c
        c(input$rankslider) -> c
        ggplot(hate, aes(y= reorder(full, row), x= Victims)) + geom_col(fill= c) + labs(y= "Groups", x= "Number of Hate Crime Victims in 2018")
        
    })
  
renderPlot({modelrank()}) -> output$rank
renderPlot({modelrank2()}) -> output$rank2
```

## Notes and Disclaimer

This application ONLY accounts for the statistics garnered in the United States in the year 2018.

The term victim may refer to an individual, business/financial institution, government entity, religious organization, or society/public as a whole.

This only accounts for the single-bias incidents in which offenses are motivated by a single bias for a group/feature; thus, there are overlaps in the choices of features due to the nature of the data.

This application does not in any way intend to discriminate a group, country, and the population within that country. Information here are solely for reference.



